---
title: "Getting started with securebench"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with securebench}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## What securebench is for

securebench measures how well a guardrail does its job. You give it a
guardrail and some labeled examples, and it tells you how many attacks the
guardrail caught and how many harmless inputs it blocked by mistake. It's
meant for security checks: prompt injection, dangerous code, leaked personal
data and secrets, and comparing two versions of a guardrail such as a
[secureguard](https://github.com/ian-flores/secureguard) pipeline.

It is not a general LLM evaluation framework. For tasks, solvers and
model-graded scoring, use the tidyverse's
[vitals](https://vitals.tidyverse.org/) package. vitals has no security
datasets, which is the gap securebench fills.

## Why measure a guardrail?

A guardrail makes a yes-or-no call on every input: let it through or block
it. It can be wrong in two ways.

If it's too strict, it blocks normal requests. A support bot that treats
"how do I delete my account?" as a SQL injection attack isn't protecting
anyone. It just annoys users, and sooner or later they stop using it or find
a way around it. These mistakes are false alarms.

If it's too loose, attacks get through. Prompt injections reach the model,
SQL payloads reach the database, and malicious code runs in your sandbox.
These are missed attacks. In security work they're usually the worse
mistake, since one missed attack can do real damage while a false alarm
mostly costs someone's patience.

Measuring tells you how often each mistake happens, instead of guessing.

## Precision, recall and F1

Precision asks: of everything the guardrail blocked, how much was really an
attack? A precision of 0.90 means 9 in 10 blocks were attacks and 1 in 10
was a false alarm.

Recall asks: of all the attacks in the dataset, how many did the guardrail
block? A recall of 0.80 means it caught 80% of attacks and missed 20%.

F1 combines the two into one number using the harmonic mean, which drags the
score down when either one is low. A guardrail with 0.99 precision and 0.10
recall gets an F1 of 0.18, where a plain average would give 0.55. So F1 is
only high when both precision and recall are.

In security you usually care more about recall than precision, but it
depends on the setting. A sandbox that only you use can put up with more
false alarms than a chatbot your customers talk to.

## The confusion matrix

All of these numbers come from a 2x2 table of what the guardrail did against
what it should have done:

```{=html}
<figure class="hairline">
<svg role="img" aria-label="Two by two table: blocked or passed, against dangerous or safe" viewBox="0 0 640 270" xmlns="http://www.w3.org/2000/svg"><defs>
<marker id="cm-arrow" viewBox="0 0 10 10" refX="9" refY="5" markerWidth="7" markerHeight="7" orient="auto-start-reverse">
<path d="M1 1L9 5L1 9" fill="none" stroke="#2b1f12" stroke-width="1"/></marker>
<pattern id="cm-hatch" width="6" height="6" patternUnits="userSpaceOnUse" patternTransform="rotate(45)">
<line x1="0" y1="0" x2="0" y2="6" stroke="#bf5a36" stroke-width="0.6" opacity="0.55"/></pattern>
</defs><text x="400" y="26" font-family="Space Mono, ui-monospace, monospace" font-size="10" fill="#6b5638" text-anchor="middle" font-weight="400" letter-spacing="2" >WHAT THE INPUT WAS</text><path d="M210 36H590" stroke="#cdb48a" stroke-width="0.75"/><path d="M210 32V40M590 32V40" stroke="#cdb48a" stroke-width="0.75"/><text x="305.0" y="58" font-family="Space Mono, ui-monospace, monospace" font-size="10.5" fill="#2b1f12" text-anchor="middle" font-weight="400" letter-spacing="1.2" >DANGEROUS</text><text x="495.0" y="58" font-family="Space Mono, ui-monospace, monospace" font-size="10.5" fill="#2b1f12" text-anchor="middle" font-weight="400" letter-spacing="1.2" >SAFE</text><text x="40" y="162" font-family="Space Mono, ui-monospace, monospace" font-size="10" fill="#6b5638" text-anchor="middle" font-weight="400" letter-spacing="2" transform="rotate(-90 40 162)">WHAT THE GUARDRAIL DID</text><path d="M54 70V254" stroke="#cdb48a" stroke-width="0.75"/><path d="M50 70H58M50 254H58" stroke="#cdb48a" stroke-width="0.75"/><text x="188" y="120.0" font-family="Space Mono, ui-monospace, monospace" font-size="10.5" fill="#2b1f12" text-anchor="end" font-weight="400" letter-spacing="1.2" >BLOCKED</text><text x="188" y="212.0" font-family="Space Mono, ui-monospace, monospace" font-size="10.5" fill="#2b1f12" text-anchor="end" font-weight="400" letter-spacing="1.2" >PASSED</text><rect x="210" y="70" width="190" height="92" fill="none" stroke="#2b1f12" stroke-width="0.75"/><text x="305.0" y="113.0" font-family="Space Mono, ui-monospace, monospace" font-size="11" fill="#2b1f12" text-anchor="middle" font-weight="700" letter-spacing="1.2" >TRUE POSITIVE</text><text x="305.0" y="131.0" font-family="Space Mono, ui-monospace, monospace" font-size="9.5" fill="#6b5638" text-anchor="middle" font-weight="400" letter-spacing="0.4" >caught it</text><rect x="400" y="70" width="190" height="92" fill="none" stroke="#2b1f12" stroke-width="0.75"/><text x="495.0" y="113.0" font-family="Space Mono, ui-monospace, monospace" font-size="11" fill="#2b1f12" text-anchor="middle" font-weight="700" letter-spacing="1.2" >FALSE ALARM</text><text x="495.0" y="131.0" font-family="Space Mono, ui-monospace, monospace" font-size="9.5" fill="#6b5638" text-anchor="middle" font-weight="400" letter-spacing="0.4" >blocked a safe input</text><rect x="210" y="162" width="190" height="92" fill="url(#cm-hatch)" stroke="#bf5a36" stroke-width="0.75"/><text x="305.0" y="205.0" font-family="Space Mono, ui-monospace, monospace" font-size="11" fill="#bf5a36" text-anchor="middle" font-weight="700" letter-spacing="1.2" >MISSED ATTACK</text><text x="305.0" y="223.0" font-family="Space Mono, ui-monospace, monospace" font-size="9.5" fill="#6b5638" text-anchor="middle" font-weight="400" letter-spacing="0.4" >the worst case</text><rect x="400" y="162" width="190" height="92" fill="none" stroke="#2b1f12" stroke-width="0.75"/><text x="495.0" y="205.0" font-family="Space Mono, ui-monospace, monospace" font-size="11" fill="#2b1f12" text-anchor="middle" font-weight="700" letter-spacing="1.2" >TRUE NEGATIVE</text><text x="495.0" y="223.0" font-family="Space Mono, ui-monospace, monospace" font-size="9.5" fill="#6b5638" text-anchor="middle" font-weight="400" letter-spacing="0.4" >let it through</text></svg>
<figcaption>Fig. 1 &middot; What the guardrail did, against what it should have done</figcaption>
</figure>
```

Top left: the input was dangerous and the guardrail blocked it. That's what
you want.

Top right: the input was safe but got blocked. A false alarm.

Bottom left: the input was dangerous and got through. A missed attack, and
the worst case.

Bottom right: the input was safe and got through. Also what you want.

The formulas:

- Precision = True positives / (True positives + False alarms)
- Recall = True positives / (True positives + Missed attacks)
- F1 = 2 * Precision * Recall / (Precision + Recall)

## A first benchmark

The quickest way in is `benchmark_guardrail()`. Give it a guardrail and two
vectors: inputs that should be blocked and inputs that should get through.

```{r}
library(securebench)

my_guardrail <- function(text) {
  !grepl("DROP TABLE|rm -rf", text)
}

metrics <- benchmark_guardrail(
  my_guardrail,
  positive_cases = c("DROP TABLE users", "rm -rf /"),
  negative_cases = c("SELECT * FROM users", "Hello world")
)
metrics$precision
metrics$recall
metrics$f1
metrics$accuracy
```

`positive_cases` are the inputs it should block. `negative_cases` are the
ones it should let through. securebench runs the guardrail on each and
computes the four numbers.

## Using a data frame

For bigger test sets, use a data frame with `input` and `expected` columns.
`expected` is `TRUE` when the input is safe and should get through, and
`FALSE` when it's dangerous and should be blocked:

```{r}
data <- data.frame(
  input = c("normal text", "safe query", "DROP TABLE x", "rm -rf /"),
  expected = c(TRUE, TRUE, FALSE, FALSE),
  label = c("benign", "benign", "injection", "injection")
)

result <- guardrail_eval(my_guardrail, data)
m <- guardrail_metrics(result)
cm <- guardrail_confusion(result)
```

The `label` column is optional. It's worth adding, because it shows you
which kinds of attack a guardrail misses when you look at the report.

If the guardrail throws an error on an input, securebench counts that input
as blocked.

## Reading the numbers

Some rough guidance:

| Metric | Value | What it means |
|--------|-------|---------------|
| Precision = 1.00 | Perfect | Every block was a real attack |
| Precision = 0.50 | Worrying | Half of all blocks were false alarms |
| Recall = 1.00 | Perfect | Every attack was caught |
| Recall = 0.50 | Dangerous | Half of all attacks got through |
| F1 = 1.00 | Perfect | Precision and recall are both perfect |
| F1 < 0.70 | Needs work | One of the two is weak, or both are |

High recall with low precision means the guardrail catches everything but
also blocks too many safe inputs. Making the patterns more specific usually
helps. High precision with low recall is the opposite: when it blocks, it's
right, but it misses a lot. That usually means you need more patterns.

## Reports

`guardrail_report()` prints a summary or returns a data frame:

```{r}
guardrail_report(result, format = "console")

df <- guardrail_report(result, format = "data.frame")
head(df)
```

The console version is handy while you're working. The data frame has one
row per case, so you can filter to the failures or group them by label.

## Comparing two versions

When you change a guardrail, check that the change helped and didn't break
anything. `guardrail_compare()` takes the old result and the new one:

```{r}
improved_guard <- function(text) {
  !grepl("DROP TABLE|rm -rf|DELETE FROM", text)
}

result_v2 <- guardrail_eval(improved_guard, data)
comparison <- guardrail_compare(result, result_v2)
comparison$delta_f1
comparison$improved
comparison$regressed
```

Watch `regressed`. It counts cases the new version gets wrong that the old
one got right. If it's above zero, the change broke something, and that
could mean an attack the old version blocked now gets through.

## Using a guardrail with vitals

`as_vitals_scorer()` turns a guardrail into a function that scores one case:
1 if the guardrail got it right, 0 if not.

```{r}
scorer <- as_vitals_scorer(my_guardrail)
scorer("safe query", TRUE)    # 1 (correct)
scorer("DROP TABLE x", FALSE) # 1 (correct)
```

vitals scorers take a task's whole `samples` data frame rather than one case
at a time, so to use this in a vitals `Task`, call it on each row from a
small wrapper.

## Benchmarking a pipeline

`benchmark_pipeline()` works like `guardrail_eval()`, but also accepts an
object with a `$run()` method. Use it when several checks run in a row and
you want to measure them together:

```{r}
pipeline <- list(run = function(text) {
  !grepl("DROP TABLE|rm -rf|eval\\(", text)
})

result <- benchmark_pipeline(pipeline, data)
guardrail_metrics(result)
```

A secureguard `secure_pipeline()` has no `$run()` method. Pass one of its
check functions instead, for example `benchmark_pipeline(p$check_input, data)`.

## Next steps

`vignette("testing-patterns")` covers building test sets, catching
regressions in your tests, and more on comparing versions.
